Methodology for Assessing the State of the Practice for Domain X

### 2020-12-04 StateOfPractice/Methodology/Methodology.tex
Further work on the introduction section of the methodology

>>>>

Research software uses computing to simulate mathematical models of real world
systems so that we can better understand and predict the system's behaviour.  A
small sample of some important examples of research software include the
following: designing new automotive parts, analysing the flow of blood in the
body, and determining the concentration of a pollutant released into the
groundwater.  As these examples illustrate, research software can be used for
tackling important problems that impact such areas as manufacturing, financial
planning, environmental policy, and the health, welfare and safety of
communities.

Given the importance of research software, scientists and engineers are pushing
for sustainable development practices.  This is evident from the existence of
such groups as SEI, BME.  Groups promoting sustainable practices are necessary
because the current state-of-the-practice for research software does not
incorporate state-of-the-art SE tools and methods
\citep{JohansonAndHasselbring2018}.  The lack of SE tools and methods
contributes to sustainability and reliability problems \citep{FaulkEtAl2009}.
Problems with the current state of the practice are evident from embarrassing
failures, like a retraction of derived molecular protein structures
\citep{Miller2006}, false reproduction of sonoluminescent fusion
\citep{PostAndVotta2005}, and fixing and then reintroducing the same error in a
large code base three times in 20 year \citep{MilewiczAndRaybourn2018}.  To
improve this situation, we need to more fully understand the current state of
the practice for research software.

To understand the current quality of SCS and to assist in developing methods for
measuring qualities, we will identify several domains within SCS and endeavour
to summarize the current state of the practice within those domains.  This
``state of the practice'' exercise will build off of prior work on
measuring/assessing the state of software development practice in several SCS
domains.  We will update the work that was done previously for domains such as
Geographic Information Systems \citep{SmithEtAl2018_arXivGIS}, Mesh Generators
\citep{SmithEtAl2016}, Seismology software \citep{SmithEtAl2018}, and
Statistical software for psychology \citep{SmithEtAl2018_StatSoft}.  We may
return to these domains and/or introduce new domains.  Potential
domains/collaborators are listed in the subsections below.

In the previous project, we measured 30 software projects for each domain, but
the measures were relatively shallow.  With the re-boot we will refine the llist
into a short list of the best projects for more detailed measurement.

In the previous state of the practice assessing exercise
\citep{SmithEtAl2018_arXivGIS}, a series of questions and simple metrics was
devised to measure the quality of the documentation and adherence to best
software development practices.  The previous set of questions is provided
\href{https://github.com/adamlazz/DomainX/blob/master/TemplateToGradeSCSoft.pdf}
{here}. As described below, the new project critically assess the previous set of questions and
revises as necessary.  In addition, the following data is also collected for each domain:

\begin{itemize}
\item Characterization of the functionality provided by the software in the
  domain.  What services does each member of the domain provide?  This
  information would be summarized in a commonality analysis document.
\item Usability testing of each software in the domain.
\item Collect empirical software engineering related data, such as the number of
  files, number of lines of code, number of open issues,
  number of closed issues, the number of lines of documentation (?). 
\end{itemize}

One of the guidelines for successfully on-boarding new developers is to
  ``Keep knowledge up to date and findable'' \citep{ShollerEtAl2019}.  The
  selected measurements should be able to assist with judging whether a project
  succeeds on this front.

The above data will be combined to rank the software in the domain.  The
specifics are yet to be determined, but in the past the Analytic Hierarchy
Process (AHP) \citep{Saaty1980} proved helpful in this.  In the past the AHP
results were calculated by a grade out of 10 from each student evaluator for
each quality for each measured product.  A better approach may be to have a
group of people familiar with the software do a pairwise comparison between each
product to come up with a number.  The number would be based on the data
collected, but not in a numerical way.  The data collected would serve the same
role as a decision support system.  It would provided insight, but human beings
would do the final pairwise comparisons, and thus the final ranking.

Key to the success of this exercise will be to involve and engage a partner for
each state of practice project.  In the previous effort we did not involve domain
experts with the rationale of excluding potential bias, but this advantage is
not worth not being able to evaluate the functionality/usability of the software.
Moreover, not having an expert makes publication more difficult, since there is
no one to advice on how best to approach journals and publishers.  The domain
expert will be asked to help in the following ways:

GitHub will be used to coordinate the work of the large team of people that will
be involved in this project.  In addition to the project record left on GitHub,
the final data will be exported to Mendeley (as was done for the previous
quality measurement exercise, for instance the grades for GIS software are
summarized at \href{https://data.mendeley.com/datasets/6kprpvv7r7/1}
{https://data.mendeley.com/datasets/6kprpvv7r7/1}.)

The data for each of the different subdomains within SCS will be standardized to
facilitate a meta-analysis that combines all of the data.
